Racial and Ethnic Differences in ETI: Outcomes

### 2024-07-09 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
started flowchart

```{r data import}
# Run code for base dataset (takes forever)
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
# Import
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
```

@@

```{r filtering}
# Additional quick cleaning steps
# Remove impossible lung function values
encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
  encounter$gli_fev1_ppred_rn > 150] <- NA
encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 |
  encounter$gli_fvc_ppred_rn > 150] <- NA
encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 |
  encounter$gli_fev1fvc_ppred_rn > 150] <- NA
# Filter
encounter <- encounter %>%
  filter(
    !is.na(Days),
    Days >= -365.25 * 3 & Days <= 365.25 * 2,
    hospitalized == "No", !encounterlocation %in% c("Hospital", "Home IV"),
    pregnant != 1,
    !is.na(Modulator_trikafta_first_date)
  )
# Get some summary statistics by participant
t1_participant <- encounter %>%
  group_by(eDWID) %>%
  summarise(
    `Years With Data` = length(unique(reviewyear)),
    `Number of Encounters` = n(),
    `Number of Encounters Pre-ETI` = sum(Days <= 0),
    `Number of Encounters Post-ETI` = sum(Days > 0),
    Age = mean(encounterage, na.rm = T),
    across(all_of(continuous_outcomes), \(x) mean(x, na.rm = TRUE)),
    Race = names(sort(table(Race), decreasing = TRUE))[1],
    Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
    .groups = "drop"
  )
no_post <- t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
encounter <- encounter %>% filter(!eDWID %in% no_post)
t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
```

@@

```{r functions}
# Plots
loess_plot <- function(data = encounter, outcome, facet = F) {
  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
    geom_point(
      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
      alpha = 0.1
    ) +
    geom_smooth(method = "loess",se = F) +
    theme_classic()
  if (facet) {
    p <- p + facet_wrap(~Race)
  }
  return(p)
}
```

>>>>

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
```

```{r filtering and flowchart}
# Variables for flowchart
n_enc_1 <- nrow(encounter)
n_people_1 <- length(unique(encounter$eDWID))
n_hosp_1 <- nrow(hospitalizations)
n_people_hosp_1 <- length(unique(hospitalizations$eDWID))
# Lung function cutoffs
lower <- 20
upper <- 150
# Remove impossible lung function values
low_fev1 = which(encounter$gli_fev1_ppred_rn)
encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
  encounter$gli_fev1_ppred_rn > 150] <- NA
encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 |
  encounter$gli_fvc_ppred_rn > 150] <- NA
encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 |
  encounter$gli_fev1fvc_ppred_rn > 150] <- NA
# Filter
encounter <- encounter %>%
  filter(
    !is.na(Days),
    Days >= -365.25 * 3 & Days <= 365.25 * 2,
    hospitalized == "No", !encounterlocation %in% c("Hospital", "Home IV"),
    pregnant != 1,
    !is.na(Modulator_trikafta_first_date)
  )
# Get some summary statistics by participant
t1_participant <- encounter %>%
  group_by(eDWID) %>%
  summarise(
    `Years With Data` = length(unique(reviewyear)),
    `Number of Encounters` = n(),
    `Number of Encounters Pre-ETI` = sum(Days <= 0),
    `Number of Encounters Post-ETI` = sum(Days > 0),
    Age = mean(encounterage, na.rm = T),
    across(all_of(continuous_outcomes), \(x) mean(x, na.rm = TRUE)),
    Race = names(sort(table(Race), decreasing = TRUE))[1],
    Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
    .groups = "drop"
  )
no_post <- t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
encounter <- encounter %>% filter(!eDWID %in% no_post)
t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
```

@@

```{r filtering}
# Additional quick cleaning steps

```

@@

```{r functions}
# Plots
loess_plot <- function(data = encounter, outcome, facet = F) {
  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
    geom_point(
      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
      alpha = 0.1
    ) +
    geom_smooth(method = "loess", se = F) +
    theme_classic()
  if (facet) {
    p <- p + facet_wrap(~Race)
  }
  return(p)
}
```
